---
title: integrateR
output: 
  github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

Thanks for this: https://r-pkgs.org/

<!-- badges: start -->
<!-- badges: end -->

Provides tools for integrated network analysis. 

- Visualize: Causal loop diagrams (CLD), Multi-level networks
- Aggregate: Find dependencies between components of causal loop diagrams based on common effects on the rest of the CLD
- Analyze: Count motifs, identify gaps, simulate baseline networks, identify virtuous and vicious circles
- Visualize results: visualize loops, suggest ties, visualize gaps

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.


## Installation

Not at this stage yet.

## Examples

```{r}
library(integrateR)
library(DiagrammeR)
```


<!-- Recoding: -->

<!-- ```{r example} -->
<!-- library(integrater) -->
<!-- ## basic example code -->
<!-- ``` -->


<!-- ```{r} -->
<!-- animal_codebook <- data.frame( -->
<!--   animal = c("elephant", "pig", "lizard"), -->
<!--   family = c("mammal","mammal","reptile")) -->
<!-- vec <- sample(c("elephant", "pig", "lizard"), 50, TRUE) -->
<!-- vec -->
<!-- ``` -->


<!-- Let's recode this: -->

<!-- ```{r} -->
<!-- rename_based_on_codebook(input = vec, -->
<!--                          codebook = animal_codebook, -->
<!--                          rawvar = "animal", -->
<!--                          codevar = "family") -->
<!-- ``` -->


