---
title: integrateR
always_allow_html: yes
output: 
  html_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- Thanks for this: https://r-pkgs.org/ -->

<!-- Must make a hex sticker at some point: https://cran.r-project.org/web/packages/hexSticker/readme/README.html -->

<!-- For packages that depend on python stuff, useful updates here: -->
<!-- https://blog.rstudio.com/2019/12/20/reticulate-1-14/ -->

<!-- "To that end, we’ve made the following changes. If the user has not explicitly instructed reticulate to use a pre-existing Python environment, then: -->
<!-- 1) reticulate will prompt the user to download and install Miniconda; -->
<!-- 2) reticulate will prepare a default r-reticulate Conda environment, using (currently) Python 3.6 and NumPy; -->
<!-- 3) When Python is initialized, reticulate will query any loaded R packages for their Python dependencies, and install those dependencies into the aforementioned r-reticulate Conda environment." -->

<!-- badges: start -->
<!-- badges: end -->

This package provides tools to analyze and visualize multi-level networks in general, and so-called integrated networks specifically.

Multi-level networks combine multiple networks in one, e.g. social-ecological networks.
Integrated networks are specific networks combining causal loop diagrams (CLDs) and social networks.

The package has three main areas: Visualization, Aggregation and Analysis.

  - Visualization: The package visualizes CLDs and two-level networks.
  
  - Aggregation: The package implements ways to aggregate CLDs to use as inputs in multi-level networks by way of finding common causes of parts of a CLD.
  
  - Analysis: The package is in many parts a R wrapper for the excellent [sesmotifanalyser](https://gitlab.com/t.seppelt/sesmotifanalyser) Python framework written by Tim Seppelt to count multi-level network motifs, compare them to a baseline and much more. It further identifies and visualizes functional gaps in integrated networks and virtuous and vicious cycles in CLDs.


## Installation

The package is currently at a very early stage of development. Explore at your own risk and please report any issues using the [issue tracker on github](https://github.com/marioangst/integrateR/issues). You can install the package from github, using devtools:

```{r eval=FALSE}
devtools::install_github("marioangst/integrater")
```


## Examples


```{r}
library(integrateR)
```

Visualize a causal loop diagram, based on DiagrammeR. This is based on an expert interview conducted in a Swiss wetland.

```{r, warning=FALSE, echo=FALSE}
cld_viz <- visualize_cld_graph(concepts_df = cld_concepts, 
                               links_df = cld_el, 
                               id_column = "concept", 
                               label_column = "english",
                               layout = "dot",
                               color_mapping = list(Target = "gray"), 
                               shape_mapping = list(Activity = "square"))
cld_viz$plot
```



