---
title: integrateR
output: 
  github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

Thanks for this: https://r-pkgs.org/

Must make a hex sticker at some point: https://cran.r-project.org/web/packages/hexSticker/readme/README.html

For packages that depend on python stuff, useful updates here:
https://blog.rstudio.com/2019/12/20/reticulate-1-14/

"To that end, we’ve made the following changes. If the user has not explicitly instructed reticulate to use a pre-existing Python environment, then:
1) reticulate will prompt the user to download and install Miniconda;
2) reticulate will prepare a default r-reticulate Conda environment, using (currently) Python 3.6 and NumPy;
3) When Python is initialized, reticulate will query any loaded R packages for their Python dependencies, and install those dependencies into the aforementioned r-reticulate Conda environment."

<!-- badges: start -->
<!-- badges: end -->

Provides tools for integrated network analysis. 

- Visualize: Causal loop diagrams (CLD), Multi-level networks
- Aggregate: Find dependencies between components of causal loop diagrams based on common effects on the rest of the CLD
- Analyze: Count motifs, identify gaps, simulate baseline networks, identify virtuous and vicious circles
- Visualize results: visualize loops, suggest ties, visualize gaps

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.


## Installation

Not at this stage yet.

## Examples

```{r}
library(integrateR)
library(DiagrammeR)
```


<!-- Recoding: -->

<!-- ```{r example} -->
<!-- library(integrater) -->
<!-- ## basic example code -->
<!-- ``` -->


<!-- ```{r} -->
<!-- animal_codebook <- data.frame( -->
<!--   animal = c("elephant", "pig", "lizard"), -->
<!--   family = c("mammal","mammal","reptile")) -->
<!-- vec <- sample(c("elephant", "pig", "lizard"), 50, TRUE) -->
<!-- vec -->
<!-- ``` -->


<!-- Let's recode this: -->

<!-- ```{r} -->
<!-- rename_based_on_codebook(input = vec, -->
<!--                          codebook = animal_codebook, -->
<!--                          rawvar = "animal", -->
<!--                          codevar = "family") -->
<!-- ``` -->


